---
title: "Population genomic analyses of <i>Xylaria necrophora</i> from southern USA | Part 1: Summarizing metadata"
author: "Teddy Garcia-Aroca"
date: 'Last update: June 19th, 2023'
output:
  html_document: rmdformats::html_clean
pdf_document: default

bibliography: references.bib
link-citations: true
---

# Summarizing metadata for 175 specimes sequenced in current study

In the current study, a total of 175 specimens, including 166 *X. necrophora* (4 historical and 162 comtemporary) and 13 specimens from several *Xylaria* spp. were sequenced.

## Load packages

For this analysis, we only need the dplyr or tidyverse packages.

```{r}
library(dplyr)
```

## Load metadata
  
```{r}
metadata <- read.csv("../metadata/Xylaria_175_genomes.popdata.Updated.10.30.2022.csv", header = T)
```

This dataset should contain the 175 specimens sequenced in the current study.

```{r}
length(metadata$No)
```

Let's summarize data by country and state

```{r}
table1 <- metadata %>% 
  dplyr::group_by(Country, State, Species) %>% 
  dplyr::summarise(count = n()) 
table1
```

This is **TABLE 1** in the manuscript.

```{r}
write.csv(table1, "../output/tables/Table1.summaryOfSpecimensSequenced.csv")
```

## Extract *X. necrophora* metadata

To summarize *X. necrophora* metadata only, we extract the metadata associated with the species: 

```{r}
metadata.necrophora <- subset(metadata, Species == "Xylaria necrophora")
```

We should see a total of 166 specimens in this dataset

```{r}
length(metadata.necrophora$No)
```

Now, let's extract the contemporary *X. necrophora* specimens only, since these represent the soybean population plus a specimen collected in the island of Martinique in 2005. We call this object `metadata.necrophora.contemp`

```{r}
metadata.necrophora.contemp <- subset(metadata.necrophora, Species_Type == "X.necrophora_Contemp")
```

This object should contain 162 individuals

```{r}
length(metadata.necrophora.contemp$No)
```

Now we summarize again.

```{r}
metadata.necrophora.contemp %>% 
  dplyr::group_by(Country, State, Species) %>% 
  dplyr::summarise(count = n()) 
```

